---
title: "| RESEARCH PROTOCOL\n| \n| EUMAEUS: Evaluating Use of Methods for Adverse Event Under Surveillance (for vaccines)\n"
fontsize: 12pt
geometry: margin=1in
output:
  bookdown::html_document2:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float: yes
    number_sections: yes
    number_tables: yes
    css: "style.css"
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: xelatex
    md_extensions: +raw_attribute
    number_sections: yes
    # citation_package: natbib
    includes:
      before_body: title.tex
header-includes:
- \usepackage[numbers,sort&compress]{natbib}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{caption}
- \usepackage{rotating}
- \usepackage{multirow}
- \usepackage{mwe,tikz}
- \usepackage[percent]{overpic}
- \usepackage{enumitem}
- \usepackage{hyperref}
- \newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}
- \newcommand{\footerDate}{`r params$date`}
- \input{header.tex}
longtable: yes
mainfont: Arial
bibliography: Protocol.bib
params:
  date: '2020-09-24'
  version: 0.0.99
subtitle: 'Version: `r params$version`'
link-citations: true
csl: jamia.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
#knitr::knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage[table]{xcolor}', x, fixed = TRUE)})
library(dplyr)
options(knitr.kable.NA = "")

pdf2png <- function(path) {
  # only do the conversion for non-LaTeX output
  if (knitr::is_latex_output()) {
    return(path)
  }
  path2 <- xfun::with_ext(path, "png")
  img <- magick::image_read_pdf(path)
  magick::image_write(img, path2, format = "png")
  path2
}


latex_table_font_size <- 8

source("PrintCohortDefinitions.R")
```

# List of Abbreviations

```{r abbreviations, echo=FALSE}
abbreviations <- readr::read_delim(col_names = FALSE, delim = ";", trim_ws = TRUE, file = "
  CDM; Common data model
  IRB; Institutional review board
  OHDSI; Observational Health Data Science and Informatics
  OMOP; Observational Medical Outcomes Partnership
  PS; Propensity score
  RCT; Randomized controlled trial
")

tab <- kable(abbreviations, col.names = NULL, linesep = "", booktabs = TRUE)

if (knitr::is_latex_output()) {
  tab %>% kable_styling(latex_options = "striped", font_size = latex_table_font_size)
} else {
  tab %>% kable_styling(bootstrap_options = "striped")
}
```

# Responsible Parties

## Investigators

```{r parties, echo=FALSE}
parties <- readr::read_delim(col_names = TRUE, delim = ";", trim_ws = TRUE, file = "
  Investigator; Institution/Affiliation
  George Hripcsak; Department of Biomedical Informatics, Columbia University, New York, NY, USA
  Dani Prieto-Alhambra; Centre for Statistics in Medicine, NDORMS, University of Oxford, Oxford, UK
  Patrick B. Ryan; Observational Health Data Analytics, Janssen Research and Development, Titusville, NJ, USA
  Martijn J. Schuemie *; Observational Health Data Analytics, Janssen Research and Development, Titusville, NJ, USA
  Marc A. Suchard; Department of Biostatistics, University of California, Los Angeles, Los Angeles, CA, USA
")

tab <- kable(parties, booktabs = TRUE, linesep = "") %>% 
  column_spec(1, width = "10em") %>%
  column_spec(2, width = "30em") %>%
  footnote(general = "* Principal Investigator", general_title = "")

if (knitr::is_latex_output()) {
  tab %>% kable_styling(latex_options = "striped", font_size = latex_table_font_size)
} else {
  tab %>% kable_styling(bootstrap_options = "striped")
}
```

## Disclosures

This study is undertaken within Observational Health Data Sciences and Informatics (OHDSI), an open collaboration.
**GH** receives grant funding from the US National Institutes of Health and the US Food & Drug Administration and contracts from Janssen Research and Development.
**PBR** and **MJS** are employees of Janssen Research and Development and shareholders in John & Johnson.
**MAS** receives grant funding from the US National Institutes of Health and the US Food & Drug Administration and contracts from Janssen Research and Development and IQVIA.

# Abstract

**Background and Significance**: 

**Study Aims**: 

**Study Description**: 



# Amendments and Updates

```{r amendments, echo=FALSE}
amendments <- readr::read_delim(col_names = TRUE, delim = ";", trim_ws = TRUE, file = "
  Number; Date; Section of study protocol; Amendment or update; Reason
  None;;;; 
")

tab <- kable(amendments, booktabs = TRUE, linesep = "")

if (knitr::is_latex_output()) {
  tab %>% kable_styling(latex_options = "striped", font_size = latex_table_font_size)
} else {
  tab %>% kable_styling(bootstrap_options = "striped")
}
```

# Milestones

```{r dates, echo=FALSE}
dates <- readr::read_delim(col_names = TRUE, delim = ";", trim_ws = TRUE, file = "
  Milestone; Planned / actual date
  EU PAS Registration; 
  Start of analysis;
  End of analysis;
  Results presentation;
")

tab <- kable(dates, booktabs = TRUE, linesep = "") 

if (knitr::is_latex_output()) {
  tab %>% kable_styling(latex_options = "striped", font_size = latex_table_font_size)
} else {
  tab %>% kable_styling(bootstrap_options = "striped")
}
```

# Rationale and Background

A total of 3 COVID-19 vaccines have been approved for clinical use in Europe, and 2 in the USA [ref]. Many more are in pipeline, and at least two more have reported to date on phase 3 efficacy data. Although safe and effective based on large randomised controlled trials, COVID-19 vaccines will be subject to post-marketing safety studies, including both analyses of spontaneous reports (pharmacovigilance) as well as longitudinal analyses in the form of post-authorisation safety studies.  

The ENCEPP (European Network of Centres for Pharmacoepidemiology and Pharmacovigilance) methodological guidelines, in their 8th revision [ref], mention a few documents that set out standards for the conducting of vaccine safety studies. Specific aspects related to vaccine safety research are discussed in detail in different materials, including the Report of the CIOMS/WHO Working Group on Definition and Application of Terms for Vaccine Pharmacovigilance (2012), the CIOMS Guide to Active Vaccine Safety Surveillance (2017), the CIOMS Guide to Vaccine Safety Communication (2018) , the Brighton Collaboration resources, the Module 4 (Surveillance) of the e-learning training course Vaccine Safety Basics by the World Health Organization (WHO), or the recommendations on vaccine-specific aspects of the EU pharmacovigilance system outlined in the Module P.I: Vaccines for prophylaxis against infectious diseases of the Good pharmacovigilance practices (GVP). Additionally, the Accelerated Development of VAccine beNefit-risk Collaboration in Europe (ADVANCE) project has summarized methods for vaccine safety in a bespoke report [ref] covering multiple study designs, both experimental and observational in nature. Despite this plethora of literature and guidance, there is a scarcity of methodological studies on the performance of different methods for vaccine safety [ref]. 

Given the quick and increasingly global rollout of COVID-19 vaccines internationally, it is highly likely that potential safety signals will emerge, which will need a timely but robust evaluation in ‘real world’ observational studies. It is therefore urgent that we conduct large-scale evaluations of methods for vaccine safety, similar to previous work on methods for drug safety [ref our work published in HDSR]. The results of this evaluation will help us understand how these methods will perform when applied to COVID vaccines. 

# Study Objectives

# Research Methods

## Exposure-outcome pairs

### Exposures

The evaluation will center on four existing (groups of) vaccines, for specific time periods (start date to end date): 

```{r exposures-of-interest, echo=FALSE}
eois <- readr::read_csv(system.file("settings", "ExposuresOfInterest.csv", package = "Eumaeus"), col_types = readr::cols())
colnames(eois) <- SqlRender::camelCaseToTitleCase(colnames(eois))

tab <- eois %>%
  select(-1, -7, -8) %>%
  kable(booktabs = TRUE, linesep = "",
      caption = "Exposures of interest.") %>% 
  kable_styling(bootstrap_options = "striped", latex_options = "striped")

if (knitr::is_latex_output()) {
  tab %>%
    column_spec(1, width = "30em") %>%
    column_spec(2, width = "8em") %>%
    column_spec(3, width = "8em") %>%
    column_spec(4, width = "8em") %>%
    column_spec(5, width = "8em") %>%
    kable_styling(font_size = latex_table_font_size)
} else {
  tab
}
```
For some methods the period between historic start and historic end date will be used to estimate the historic incidence rate. 
The formal cohort definitions of each exposure can be found in Appendix \@ref(exposure-cohort-definitions).

### Negative control outcomes 

Negative controls are outcomes believed not to be caused by any of the vaccines, and therefore ideally would not be flagged as a signal by a safety surveillance system. 
Any effect size estimates for negative control ideally should be close to the null. 

A single set of negative control outcomes is defined for all four vaccine groups.
To identify negative control outcomes that match the severity and prevalence of suspected vaccine adverse effects, a candidate list of negative controls was generated based on similarity of prevalence and percent of diagnoses that were recorded in an inpatient setting (as a proxy for severity). 
Manual review of this list by clinical experts created the final list of 127 negative control outcomes. 
The full list of negative control outcomes can be found in Appendix \@ref(negative-controls)

Negative control outcomes are defined as any occurrence of the negative control concept or any of its descendants. 

### Positive control outcomes 

Positive controls are outcomes known to be caused by vaccines, and ideally would be detected as signals by a safety surveillance system as early as possible.
For various reasons, real positive controls are problematic.[@Schuemie2018-zi] 
Instead, here we will rely on synthetic positive controls,[@Schuemie2020-he;@Schuemie2018-hq] created by modifying a negative control through injection of additional, simulated occurrences of the outcome. 
To preserve (measured) confounding, simulated outcome occurrences are sampled from the probability distribution derived from a predictive model fitted on the data.
Target true hazard ratios for the positive control synthesis are 1.5, 2, and 4, so using the 127 negative controls we are able to construct 127 * 3 = 381 positive control outcomes.
The hazard for the outcome is simulated to be increased by the target ratio for the period starting 1 day after vaccination until 28 days after vaccinations, with a constant hazard ratio during that time. 
This increased risk is applied both for the first and second injection of multi-dose vaccines.

## Data

We will execute EUMAEUS as an OHDSI network stuy.
All data partners within OHDSI are encouraged to participate voluntarily and can do so conveniently, because of the community's shared Observational Medical Outcomes Partnership (OMOP) common data model (CDM) and OHDSI tool-stack.
Many OHDSI community data partners have already committed to participate and we will recruit further data partners through OHDSI’s standard recruitment process, which includes protocol publication on OHDSI’s GitHub, an announcement in OHDSI’s research forum, presentation at the weekly OHDSI all-hands-on meeting and direct requests to data holders.

Table \@ref(tab:data-sources) lists the 13 already committed data sources for EUMAEUS; these sources encompass a large variety of practice types and populations. 
For each data source, we report a brief description and size of the population it represents.
All data sources will receive institutional review board approval or exemption for their participation before executing EUMAEUS.

```{r data-sources, echo=FALSE, warning=FALSE}
data_sources <- readr::read_delim(col_names = TRUE, delim = ";", trim_ws = TRUE, file = "
  Data source ; Population ; Patients ; History ; Data capture process and short description
  IBM MarketScan Commercial Claims and Encounters (CCAE) ; Commercially insured, < 65 years ; 142M ; 2000 -- ; Adjudicated health insurance claims (e.g. inpatient, outpatient, and outpatient pharmacy)  from large employers and health plans who provide private healthcare coverage to employees, their spouses and dependents.
  IBM MarketScan Medicare Supplemental Database (MDCR)  ; Commercially insured, 65$+$ years ; 10M ; 2000 -- ; Adjudicated health insurance claims of retirees with primary or Medicare supplemental coverage through privately insured fee-for-service, point-of-service or capitated health plans.
  IBM MarketScan Multi-State Medicaid Database (MDCD) ; Medicaid enrollees, racially diverse ; 26M ; 2006 -- ; Adjudicated health insurance claims for Medicaid enrollees from multiple states and includes hospital discharge diagnoses, outpatient diagnoses and procedures, and outpatient pharmacy claims.
  Japan Medical Data Center (JMDC) ; Japan, general ; 5.5M ; 2005 -- ; Data from 60 society-managed health insurance plans covering workers aged 18 to 65 and their dependents.
  Optum Clinformatics Data Mart (Optum) ; Commercially or Medicare insured ; 85M ; 2000 -- ; Inpatient and outpatient healthcare insurance claims.
  Optum Electronic Health Records (OptumEHR) ; US, general ; 93M ; 2006 -- ; Clinical information, prescriptions, lab results, vital signs, body measurements, diagnoses and procedures derived from clinical notes using natural language processing. 
")
tab <- kable(data_sources, booktabs = TRUE, linesep = "",
      caption = "Committed EUMAEUS data sources and the populations they cover.") %>% 
  kable_styling(bootstrap_options = "striped", latex_options = "striped") %>%
  pack_rows("Administrative claims", 1, 5, latex_align = "c", indent = FALSE) %>%
  pack_rows("Electronic health records (EHRs)", 6, 6, latex_align = "c", indent = FALSE)

if (knitr::is_latex_output()) {
  tab %>%
    column_spec(1, width = "10em") %>%
    column_spec(2, width = "10em") %>%
    column_spec(5, width = "25em") %>%
    kable_styling(font_size = latex_table_font_size)
} else {
  tab
}
```

## Methods to evaluate

Vaccine safety surveillance methods can be broken down in to four components: construction of a *counterfactual* (often referred to as the 'expected count'), a *time-at-risk*, the *statistic* to estimate, and potentially a *decision rule* on the estimate to classify signals from non-signals.

### Counterfactual construction

- Historic rate. Traditionally, vaccine surveillance methods compute an expected count based an incidence rate estimated during some historic time period, for example in the years prior to the initiation of the surveillance study. 
    - Using a single rate for the entire population, 
    - Stratifying by age and sex, allowing the expected rate to be adjusted for the demographics of the vaccinated.
- Cohort method using a contemporary non-user comparator. We require the non-user comparator to have an outpatient visit on the index date.
    - Crude comparison
    - Using 1-on-1 propensity score matching
- Self-Controlled Case Series (SCCS) / Self-Controlled Risk Interval (SCRI).
    - A simple SCCS
    - An SCCS adjusting for age and season
    - An SCRI, using a control interval of 43 to 15 days prior to vaccination.
- Case-control
    - Using age and sex matching controls
    - By sampling controls from the general non-case population.

### Time-at-risk

The time-at-risk is the time window, relative to the vaccination date, when outcomes will potentially be attributed to the vaccine. We define three time-at-risk windows: 1-28 days, 1-42 days, and 0-1 days after vaccination. 
Time-at-risk windows will be constructed both for the first and second dose. The time-at-risk for one dose will be censored at the time of the next dose. 

### Statistic

- Effect-size estimate. Each method can be used to produce an effect-size estimates such as a hazard ratio, incidence rate ratio, or odds ratio. For example, when using a historic rate we can compute the observed to expected ratio, which can be interpreted as the incidence rate ratio.

- Log likelihood ratio (LLR). A common practice in vaccine safety surveillance is to computer the LLR, which is the log of the ratio between the likelihood of the alternative hypothesis (that there is an effect) and the likelihood of the null hypothesis (of no effect). 
The LLR is a convenient statistic when performing sequential testing, where the LLR can be compared to a pre-computed critical value, as is done in the PMaxSPRT method. [@kulldorf2011] 
Although typically PMaxSPRT uses a historic rate as counterfactual, any counterfactual can be used to compute the LLR and can be used in PMaxSPRT.

Effect-size estimates will be computed both with and without empirical calibration. [@Schuemie2014-bv;@Schuemie2018-hq]
Empirical calibration will be done using leave-one-out: when calibrating the estimate for a control, the systematic error distribution will be fitted uses all controls except the one being calibrated.

### Decision rule

To identify 'signals' we need a decision rule, for example in the shape of a threshold value on one of the estimates statistics.
In our experiment we will consider one decision rule, which is the critical value computed for the LLR at an alpha of 0.01. 
For this we will use the [`Sequential` package in CRAN](https://cran.r-project.org/web/packages/Sequential/index.html).

## Metrics

Similar to our previous study, we will compute the following metrics based on the effect size estimates: [@Schuemie2020-wx]

- Area Under the receiver-operator Curve (AUC). The ability to discriminate between positive controls and negative controls based on the point estimate of the effect size.
- Coverage. How often the true effect size is within the 95% confidence interval.
- Mean precision. Precision is computed as 1 / (standard error)2, higher precision means narrower confidence intervals. We use the geometric mean to account for the skewed distribution of the precision.
- Mean squared error (MSE). Mean squared error between the log of the effect size point-estimate and the log of the true effect size.
- Type 1 error. For negative controls, how often was the null rejected (at alpha = 0.05). This is equivalent to the false positive rate and 1 - specificity.
- Type 2 error. For positive controls, how often was the null not rejected (at alpha = 0.05). This is equivalent to the false negative rate and 1 - sensitivity.
- Non-estimable. For how many of the controls was the method unable to produce an estimate? There can be various reasons why an estimate cannot be produced, for example because there were no subjects left after propensity score matching, or because no subjects remained having the outcome.

In addition, based on the PMaxSPRT decision rule, we will compute sensitivity and specificity. 

### Timeliness

To understand the time it takes for a method the identify signals, the study period for each vaccine will be divided into calendar months. 
For each month the methods will be executed using the data that had accumulated up to the end of that month, and the performance metrics will be reported for each month.

### Multiple doses

For those vaccines requiring multiple doses (zoster, HPV), metrics will be computed three times: 

- Treating all doses the same, so computing statistics using both doses without distinguishing between first and second.
- Using the first dose only
- Using the second dose only

## Overview of analyses

In total, we will evaluate:

- 9 counterfactuals
- 3 times at risk (1-28, 1-42, and 1-90 days)
- 4 vaccines, with a total of 9 + 12 + 12 + 12 = 45 time periods
- 127 negative controls
- 3 * 127 = 381 positive controls
- 3 dose definitions (both, first, second) for the zoster and HPV vaccines, 1 for H1N1 and seasonal flu.

Resulting in a total of 9 * 3 * ((9 + 12) * 1 + (12 + 12) * 3)  * (127 + 381) = 1,275,588 effect-size estimates. Each estimate will contain:

- The effect-size estimate (e.g. hazard ratio, odds ratio) with 95% confidence interval and p-value.
- The empirically calibrated effect-size estimate and p-value
- The LLR

This will be computed for each database.

# Strengths and Limitations

## Strengths
- Cohort studies allow direct estimation of incidence rates following exposure of interest, and the new-user design can capture early events following treatment exposures while avoiding confounding from previous treatment effects; new use allows for a clear exposure index date.
- Large-scale propensity score matching and stratification create balance on a large number of baseline potential confounders and have been found in the past to balance unmeasured confounders.
- Systematic processes including a pre-specified selection of covariates avoids investigator-specific biases in variable selection.
- Use of real negative and synthetic positive control outcomes provides an independent estimate of residual bias in the experiment.
- The fully specified study protocol is being published before analysis begins.
- Dissemination of the results will not depend on estimated effects, avoiding publication bias.
- All analytic methods have previously been verified on real data.
- All software is freely available as open source.
- Use of a common data model allows extension of the experiment to future databases and allows replication of these results on licensable databases that were used in this experiment, while still maintaining patient privacy on patient-level data.
- Use of multiple databases allows estimating consistency to add credibility and supports generalizability.

 
## Limitations
 
- Even though many potential confounders will be included in this study, there may be residual bias due to unmeasured or misspecified confounders, such as confounding by indication, differences in physician characteristics that may be associated with drug choice, concomitant use of other drugs started after the index date, and informative censoring at the end of the on-treatment periods. To minimize this risk, we used methods to detect residual bias through our negative and positive controls.
- Our follow-up times are limited and variable, potentially reducing power to detect differences in effectiveness and safety.
- We assume hazards are not time varying.
- Misclassification of study variables is unavoidable in secondary use of health data, so it is possible to misclassify treatments, covariates, and outcomes; we do not expect differential misclassification, so bias will most likely be towards the null.
- The electronic health record databases may be missing care episodes for patients due to care outside the respective health systems; bias will most likely be towards the null.

# Protection of Human Subjects

EUMAEUS does not involve human subjects research.
The project does, however, use de-identified human data collected during routine healthcare provision.
All data partners executing the EUMAEUS studies within their data sources will have received institutional review board (IRB) approval or waiver for participation in accordance to their institutional governance prior to execution (see Table ADD-REF).
EUMAEUS executes across a federated and distributed data network, where analysis code is sent to participating data partners and only aggregate summary statistics are returned, with no sharing of patient-level data between organizations.

```{r irb, echo=FALSE}
data_sources <- readr::read_delim(col_names = TRUE, delim = "&", trim_ws = TRUE, file = "
Data source & Statement
IBM MarketScan Commercial Claims and Encounters (CCAE) & New England Institutional Review Board and was determined to be exempt from broad IRB approval, as this research project did not involve human subject research.
IBM MarketScan Medicare Supplemental Database (MDCR)  & New England Institutional Review Board and was determined to be exempt from broad IRB approval, as this research project did not involve human subject research.
IBM MarketScan Multi-State Medicaid Database (MDCD) & New England Institutional Review Board and was determined to be exempt from broad IRB approval, as this research project did not involve human subject research.
IQVIA Open Claims (IOC) & This is a retrospective database study on de-identified data and is deemed not human subject research. Approval is provided for OHDSI network studies.
Japan Medical Data Center (JMDC) & New England Institutional Review Board and was determined to be exempt from broad IRB approval, as this research project did not involve human subject research.
Optum Clinformatics Data Mart (Optum) & New England Institutional Review Board and was determined to be exempt from broad IRB approval, as this research project did not involve human subject research.
Columbia University Irving Medical Center (CIUMC) & Use of the CUIMC data source was approved by the Columbia University Institutional Review Board as an OHDSI network study (IRB\\# AAAO7805).
Department of Veterans Affairs (VA) & Use of the VA-OMOP data source was reviewed by the Department of Veterans Affairs Central Institutional Review Board (IRB) and was determined to meet the criteria for exemption under Exemption Category 4(3) and approved the request for Waiver of HIPAA Authorization.
Information System for Research in Primary Care (SIDIAP) & Use of the SIDIAP data source was approved by the Clinical Research Ethics Committee of IDIAPJGol (project code: 20/070-PCV)
IQVIA Disease Analyzer Germany (DAG) & This is a retrospective database study on de-identified data and is deemed not human subject research. Approval is provided for OHDSI network studies.
Optum Electronic Health Records (OptumEHR) & New England Institutional Review Board and was determined to be exempt from broad IRB approval, as this research project did not involve human subject research.
Yale New Haven Health System (YNHHS) & Use of the YNHHS EHR data source was approved by the Yale University Institutional Review Board as an OHDSI network study (IRB\\# pending).
")

tab <- kable(data_sources, booktabs = TRUE, linesep = "",
      caption = "IRB approval or waiver statement from partners.") %>% 
  kable_styling(bootstrap_options = "striped", latex_options = "striped")

if (knitr::is_latex_output()) {
  tab %>%
    column_spec(1, width = "15em") %>%
    column_spec(2, width = "40em") %>%
    kable_styling(font_size = latex_table_font_size)
} else {
  tab
}
```

# Management and Reporting of Adverse Events and Adverse Reactions

EUMAEUS uses coded data that already exist in electronic databases.
In these types of databases, it is not possible to link (i.e., identify a potential causal association between) a particular product and medical event for any specific individual.
Thus, the minimum criteria for reporting an adverse event (i.e., identifiable patient, identifiable reporter, a suspect product and event) are not available and adverse events are not reportable as individual adverse event reports.
The study results will be assessed for medically important findings.

# Plans for Disseminating and Communicating Study Results

Open science aims to make scientific research, including its data process and software, and its dissemination, through publication and presentation, accessible to all levels of an inquiring society, amateur or professional [@Woelfle2011-ss] and is a governing principle of EUMAEUS.
Open science delivers reproducible, transparent and reliable evidence.
All aspects of EUMAEUS (except private patient data) will be open and we will actively encourage other interested researchers, clinicians and patients to participate.
This differs fundamentally from traditional studies that rarely open their analytic tools or share all result artifacts, and inform the community about hard-to-verify conclusions at completion.

## Transparent and re-usable research tools

We will publicly register this protocol and announce its availability for feedback from stakeholders, the OHDSI community and within clinical professional societies.
This protocol will link to open source code for all steps to generating diagnostics, effect estimates, figures and tables.
Such transparency is possible because we will construct our studies on top of the OHDSI toolstack of open source software tools that are community developed and rigorously tested [@Schuemie2020-wx].
We will publicly host EUMAEUS source code at (https://github.com/ohdsi-studies/Eumaeus), allowing public contribution and review, and free re-use for anyone’s future research.

## Continous sharing of results

EUMAEUS embodies a new approach to generating evidence from healthcare data that overcome weaknesses in the current process of answering and publishing (or not) one question at a time.
Generating evidence for thousands of research and control questions using a systematic process enables us to not only evaluate that process and the coherence and consistency of the evidence, but also to avoid $p$-hacking and publication bias [@Schuemie2018-zi].
We will store and openly communicate all of these results as they become available using a user-friendly web-based app that serves up all descriptive statistics, study diagnostics and effect estimates for each cohort comparison and outcome.
Open access to this app will be through a general public facing EUMAEUS web-page.

## Scientific meetings and publications

We will deliver multiple presentations annually in scientific venues including the annual meetings of the American Diabetes Association, American College of Cardiology, American Heart Association and American Medical Informatics Association.
We will also prepare multiple scientific publications for clinical, informatics and statistical journals.

## General public

We believe in sharing our findings that will guide clinical care with the general public.
EUMAEUS will use social-media (Twitter) to facilitate this.
With dedicated support from the OHDSI communications specialist, we will deliver regular press releases at key project stages, distributed via the extensive media networks of UCLA, Columbia and Yale.

# Lists of Tables and Figures

# References {-}

<div id="refs"></div>

# (APPENDIX) Appendix {-}

# Exposure Cohort Definitions

```{r h1n1-cohort, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
baseCohortJson <- SqlRender::readSql(system.file("cohorts", "H1N1vaccination.json", package = "Eumaeus"))
baseCohort <- RJSONIO::fromJSON(baseCohortJson)

baseCohortJson <- RJSONIO::toJSON(baseCohort, digits = 50)

printCohortDefinitionFromNameAndJson(name = "H1N1 Vaccines",
                                     json = baseCohortJson)
```

```{r flu-cohort, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
baseCohortJson <- SqlRender::readSql(system.file("cohorts", "FluVaccination.json", package = "Eumaeus"))
baseCohort <- RJSONIO::fromJSON(baseCohortJson)

baseCohortJson <- RJSONIO::toJSON(baseCohort, digits = 50)

printCohortDefinitionFromNameAndJson(name = "Seasonal Flu Vaccines",
                                     json = baseCohortJson)
```

```{r hpv-cohort, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
baseCohortJson <- SqlRender::readSql(system.file("cohorts", "HpvVaccination.json", package = "Eumaeus"))
baseCohort <- RJSONIO::fromJSON(baseCohortJson)

baseCohortJson <- RJSONIO::toJSON(baseCohort, digits = 50)

printCohortDefinitionFromNameAndJson(name = "HPV Vaccines",
                                     json = baseCohortJson)
```

```{r zoster-cohort, echo=FALSE, results="asis", warning=FALSE, message=FALSE}
baseCohortJson <- SqlRender::readSql(system.file("cohorts", "ZosterVaccination.json", package = "Eumaeus"))
baseCohort <- RJSONIO::fromJSON(baseCohortJson)

baseCohortJson <- RJSONIO::toJSON(baseCohort, digits = 50)

printCohortDefinitionFromNameAndJson(name = "Zoster Vaccines",
                                     json = baseCohortJson)
```


# Negative controls

```{r ncs, echo=FALSE}
ncs <- readr::read_csv(system.file("settings", "NegativeControls.csv", package = "Eumaeus"), col_types = readr::cols())
colnames(ncs) <- SqlRender::camelCaseToTitleCase(colnames(ncs))

tab <- kable(ncs, booktabs = TRUE, linesep = "",
      caption = "Negative control outcomes.") %>% 
  kable_styling(bootstrap_options = "striped", latex_options = "striped")

if (knitr::is_latex_output()) {
  tab %>%
    column_spec(1, width = "15em") %>%
    column_spec(2, width = "40em") %>%
    kable_styling(font_size = latex_table_font_size)
} else {
  tab
}
```

# Analysis Plan Matrix

